Learning \strips\ action models from {\em state-invariants}

### 2019-02-05 doc/ijcai-19-deductivelearning/deductivelearning.tex
first updates to section 3 ijcai19 paper on deductive learning

We define the task of learning action models as a tuple $\Lambda=\tup{M,\Phi,\mathcal{O}(\tau)}$, where:
\begin{itemize}
\item $M$ is the {\em space of possible action model}. This set is the {\em full} space of action models, when learning from scratch, or {\em partially specified}, when some fragments of the action models are known a priori. A set of action models can be defined {\em explicitly}, enumerating all the models that belong to the set or {\em implicitly}, enumerating all the constraints that must satisfy any model that belongs to the set. A {\em partially specified} \strips\ action model is then a formalism for the {\em implicit} representation of a set of \strips\ schemes~\cite{sreedharan2018handling}. 
\item $\Phi$, a set of {\em state-invariants} that constrain the set of possible states in the given domain.
\item $\mathcal{O}(\tau)$ is an observation of a trajectory $\tau(\pi,P)$ produced by the execution of an unknown plan $\pi$ that solves $P_\mathcal{O}$ (i.e. the classical planning problem $P_\mathcal{O}=\tup{F,A[\cdot],s_0^o,s_m^o}$ where $A[\cdot]$ is a set of actions s.t., the semantics of each action $a\in A[\cdot]$ is unknown since functions $\rho$ and/or $\theta$ of $a$ are undefined). 
\end{itemize}

A {\em solution} to a $\Lambda=\tup{M,\Phi,\mathcal{O}(\tau)}$ learning task is a model $\mathcal{M}'\in M$ that is {\em consistent} with the given input knowledge $\Phi$ and $\mathcal{O}(\tau)$. This means that let there exists a plan $\pi$ such that the observation $\mathcal{O}(\tau)$ is consistent with $\pi$ when the dynamics of $A[\cdot]$ is given by $\mathcal{M}'\in M$ and such that any state produced in the trajectory $\tau(\pi,P_\mathcal{O})$ satisfies the {\em state-invariants} $\Phi$.

%They compute an action model that maximizes some notion of {\em statistical consistency} over a set of observations of plan executions so output an action model that is {\em consistent} with the input knowledge.

@@

\subsection{A propositional encoding for the space of STRIPS action models}
{\em A \strips\ action schema} $\xi$ is defined by four lists: A list of {\em parameters} $pars(\xi)$, and three list of predicates (namely $pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schema. Let be $\Psi$ the set of {\em predicates} that shape the propositional state variables $F$, and a list of {\em parameters} $pars(\xi)$. The set of elements that can appear in $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of the \strips\ action schema $\xi$ is given by FOL interpretations of $\Psi$ over the parameters $pars(\xi)$ and is denoted as ${\mathcal I}_{\Psi,\xi}$.

>>>>

We define the task of learning an action model from {\em state-invaraints} as a tuple $\Lambda=\tup{F,I,G,\Phi,M}$, where:
\begin{itemize}
\item $F$ is a state of fluents, $I$ is an initial state (a full assigment of the fluents), and $G$ is a set of goal conditions over the fluents.
\item $\Phi$, a set of {\em state-invariants} that constrain the set of possible states in the given domain.
\item $M$ is the {\em space of possible action model}. This set is the {\em full} space of action models, when learning from scratch, or {\em partially specified}, when some fragments of the action models are known a priori. A set of action models can be defined {\em explicitly}, enumerating all the models that belong to the set or {\em implicitly}, enumerating all the constraints that must satisfy any model that belongs to the set. A {\em partially specified} \strips\ action model is then a formalism for the {\em implicit} representation of a set of \strips\ schemes~\cite{sreedharan2018handling}.
\end{itemize}

A {\em solution} to a $\Lambda=\tup{F,I,G,\Phi,M}$ learning task is a model $\mathcal{M}'\in M$ that is {\em consistent} with the given input knowledge. This means that, let $P=\tup{F,A_{\mathcal{M}'},I,G}$ be a {\em classical planning problem} where the dynamics of the actions is given by the action model $\mathcal{M}'$, there exists a plan $\pi$ such that any state traversed in the trajectory $\tau(\pi,P_\mathcal{O})$ satisfies all the {\em state-invariants} in $\Phi$.

@@

\subsection{A propositional encoding for the space of STRIPS action models}
%They compute an action model that maximizes some notion of {\em statistical consistency} over a set of observations of plan executions so output an action model that is {\em consistent} with the input knowledge.

{\em A \strips\ action schema} $\xi$ is defined by four lists: A list of {\em parameters} $pars(\xi)$, and three list of predicates (namely $pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schema. Let be $\Psi$ the set of {\em predicates} that shape the propositional state variables $F$, and a list of {\em parameters} $pars(\xi)$. The set of elements that can appear in $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of the \strips\ action schema $\xi$ is given by FOL interpretations of $\Psi$ over the parameters $pars(\xi)$ and is denoted as ${\mathcal I}_{\Psi,\xi}$.
